Beyond Ultra-Diffuse Galaxy: Mass-Size Outliers in the Satellites of Milky Way Analogs

### 2022-09-13 main.tex
before co-author's comments

\author[0000-0002-8873-5065]{Peter Melchior}
\affiliation{Department of Astrophysical Sciences, 4 Ivy Lane, Princeton University, Princeton, NJ 08544, USA}
\author[0000-0002-2704-5028]{Remy Joseph}
\affiliation{Department of Astrophysical Sciences, 4 Ivy Lane, Princeton University, Princeton, NJ 08544, USA}
\author[0000-0002-1841-2252]{Rachael Beaton}
\affiliation{Department of Astrophysical Sciences, 4 Ivy Lane, Princeton University, Princeton, NJ 08544, USA}
\author[0000-0002-2991-9251]{Kirsten Casey}
\affiliation{Center for Cosmology and AstroParticle Physics (CCAPP), The Ohio State University, Columbus, OH 43210, USA}
\author[0000-0002-1841-2252]{Shany Danieli}
\affiliation{Department of Astrophysical Sciences, 4 Ivy Lane, Princeton University, Princeton, NJ 08544, USA}
\author[0000-0002-1841-2252]{Andy Goulding}
\affiliation{Department of Astrophysical Sciences, 4 Ivy Lane, Princeton University, Princeton, NJ 08544, USA}
\author[0000-0002-1841-2252]{Erin Kado-Fong}
\affiliation{Department of Astrophysical Sciences, 4 Ivy Lane, Princeton University, Princeton, NJ 08544, USA}

@@

%% Mark off the abstract in the ``abstract'' environment. 
\begin{abstract}
\jiaxuan{Need to rewrite abstract}
Ultra-diffuse galaxies (UDGs) are dwarf galaxies with extreme sizes and low surface brightness. The formation and quenching mechanisms of these intriguing objects are still under debate. In this work, we present systematic search for ultra-diffuse galaxies associated with Milky Way (MW) analogs in the nearby Universe ($0.01 < z < 0.04$) using the Hyper Suprime-Cam Strategic Survey Program data. We also define a new subset of low surface brightness galaxies dubbed as ``ultra-puffy galaxies'' (UPGs) which better characterize the size outliers on the mass-size plane. We perform statistical studies for UDGs and UPGs and compare the results with previous studies for UDGs in clusters and groups. We further contrast UDGs and UPGs with the bulk of MW satellites in nearby Universe and simulations. We find that UDGs have a higher quenched fraction than MW satellites and the quenched fraction remains constant over 1 dex in stellar mass. We argue that this high quiescent fraction is an artifact of the UDG definition. Despite their large sizes, UPGs in MW analogs have similar quenched fractions as the bulk satellites. These results hint that the size growth and quenching of UPGs must happen separately. 
\end{abstract}

@@

\section{Introduction} \label{sec:intro}
In a variety of galaxy populations, an interesting subset of galaxies with low surface brightness and large physical size have been identified in the past \citep[dubbed as LSBGs, e.g.,][]{Sandage1984,Caldwell1987,Impey1988,McGaugh1995,Dalcanton1997a}. Recently, new excitement was generated by the discovery by \citet{vanDokkum2015} of a large population of ultra-diffuse galaxies (UDGs) in the Coma cluster, accompanied by many other works on detecting UDGs in clusters \citep[e.g.,][]{Koda2015,Mihos2015,Yagi2016,vdBurg2016,vdBurg2017,Lee2017,ManceraPina2018,Zaritsky2019}, groups \citep[e.g.,][]{Roman2017b,Greco2018,SAGA-II,CarlstenELVES2022}, and fields \citep[e.g.,][]{Leisman2017,Roman2019,Prole2019,Tanoglidis2021,Kadowaki2021}. There have been a variety of studies of these extremely-diffuse galaxy systems, ranging from their dark matter content \citep[e.g.,][]{Mowla2017,vanDokkum2018,vanDokkum2019,Wasserman2019,Keim2022}, to their globular cluster populations \citep[e.g.,][]{vanDokkum2017,Somalwar2020,Forbes2020,Danieli2022,Gannon2022}, and stellar populations \citep[e.g.,][]{Gu2018,Ferre-Mateu2018,Villaume2022}. The UDGs are dwarf galaxies with larger size ($r_e > 1.5$ kpc) and lower surface brightness ($\sbcen > 24.0\ \sbunit$) than normal dwarfs, thus are outliers from the average mass-size relation. Therefore, a central question about mass-size outliers including UDGs is what physical mechanisms are responsible for the formation and quenching of these large diffuse galaxies.

% In \citet{vanDokkum2015}, the ultra-diffuse galaxies (UDGs) are defined as dwarf galaxies with large size ($r_e > 1.5$ kpc) and low surface brightness ($\sbcen > 24.0\ \sbunit$).

Different mechanisms are proposed to explain the presence of large diffuse galaxies in different environments. Field UDGs might form after star-formation feedback \citep{DiCintio2017,Chan2018}, early galaxy mergers \citep{Wright2021}, passing into and out of a more massive halo (so-called backsplash satellites, \citealt{Benavides2021}), or they were born in halos with higher spin \citep{Dalcanton1997,Amorisco2016,Liao2019}. UDGs in groups and clusters are believed to be puffed up by tidal heating near pericenter \citep{Jiang2019} or to undergo adiabatic expansion due to mass loss \citep{Tremmel2020}. Despite theoretical advances, it is still an open question how mass-size outliers like UDGs in certain environments are formed and quenched. It is also unclear whether mass-size outliers belong to a special class of galaxy or are just a natural extension of dwarf systems to lower surface brightness. To answer these questions, we can compare mass-size outliers with the ``normal'' satellites which we have a better understanding.

Our best knowledge in the dwarf galaxy regime is about the satellite system of our Milky Way (MW) and the Local Group (LG) \citep[e.g.,][]{McConnachie2012,Simon2019}. With the advent of deep sky surveys and dedicated spectroscopy programs, we are now mapping the satellite systems of MW analogs in the Local Volume and nearby Universe. The Satellites Around Galactic Analogs (SAGA) survey is designed to spectroscopically confirm the satellites of 100 MW analogs at $z\sim0.01$ \citep{SAGA-I,SAGA-II}. The Exploration of Local VolumE Satellites (ELVES) survey uses deep imaging data to identify satellite candidates of MW-like hosts in the Local Volume (LV, $D<12$ Mpc) and estimate surface brightness fluctuation distances from images to confirm satellites \citep{ELVES-I,ELVES-II,CarlstenELVES2022}. Having the MW satellites as a reference frame, we can study the formation and quenching of large diffuse galaxies by comparing them with the bulk of observed and simulated MW satellites. However, there are only $\sim 30$ confirmed UDGs associated with MW analogs in the literature \citep{Roman2017b,Cohen2018,SAGA-II,CarlstenELVES2022}. A larger sample of mass-size outliers (including UDGs) associated with MW hosts is needed to conduct such studies.

The goal of this paper is to study large diffuse galaxies in the reference frame of the satellite systems of MW analogs and try to elucidate their formation and quenching. Albeit large size and low surface brightness, UDGs are not necessarily mass-size outliers in terms of their offsets to the average mass-size relation (see \S \ref{sec:sample_construction}). Motivated by this fact, we propose studying ``ultra-puffy galaxies'' (UPGs) which are $1.5\sigma$ above the average mass-size relation, representing the large-size tail of the dwarf galaxy population. Using the Hyper-Suprime Camera imaging data, we construct a sample of low surface brightness galaxies (LSBGs) and match them with MW analogs at $0.01 < z < 0.04$. Assuming the host distance, we calculate the physical size and stellar mass for LSBGs and further construct the UDG and UPG samples. We compare the statistical properties of mass-size outliers (UDGs and UPGs) with LV satellites and simulations to shed light on questions about their formation and evolution.

@@

% UDGs found in groups and clusters are mostly red and quenched, whereas their blue counterparts are found in field environments (although several quenched UDGs are found in the field, \citealt{Roman2019,Prole2019}). This suggests that environmental effects play a critical role in the formation and quenching of UDGs.

The layout of this paper is as follows. Section \ref{sec:data} describes the data used in this work. In Section \ref{sec:lsbg_search}, we describe how we search the LSBG candidates in HSC data, including source detection (\S \ref{sec:detection}), deblending (\S \ref{sec:deblending}), and modeling (\S \ref{sec:modeling}). In Section \ref{sec:sample_construction}, we propose the UPG definition and present our UDG and UPG samples. In Section \ref{sec:results}, we show statistical analysis for UDGs and UPGs including their abundances (\S \ref{sec:n_udg}), size distributions (\S \ref{sec:size_distr}), spatial distributions (\S \ref{sec:radial_distr}), and quenched fractions (\S \ref{sec:quench}). We argue why we need to move beyond UDG and discuss possible implications of our results on the formation and quenching of mass-size outliers in Section \ref{sec:discussion}. Section \ref{sec:summary} presents a summary of this work and prospects for the future. In this work, we refer to LSBGs as objects selected purely by surface brightness, whereas UDGs and UPGs are selected by both physical size and surface brightness. We adopt a flat $\Lambda$CDM cosmology from \citet{Planck15} with $\Omega_{\rm m}= 0.307$ and $H_0 = 67.7\ $km s$^{-1}$ Mpc$^{-1}$. We use the AB system \citep{Oke1983} for magnitudes. The stellar mass used in this work is based on a \citet{Chabrier2003} initial mass function.

% Continuing the work by \citep{Greco2018}, we conducted a systematic search for low surface brightness galaxies (LSBGs) in Hyper Suprime-Cam survey (HSC) imaging data (Li et al., in prep). In order to study the formation and evolution of UDGs in low-density environments, we first select host galaxies at $0.01 < z < 0.04$ with Milky-Way masses ($10.2 < \log M_*/M_\odot < 11.2$) in the NASA-Sloan Atlas. We matched our HSC LSBGs catalog with the MW-like hosts catalog. We assign an LSBG to a host if their angular separation is smaller than the projected virial radius of the host. Then we select a UDG sample by requiring the $g$-band average surface brightness within the effective radius\footnote{This surface brightness threshold is equivalent to $\mu_0(g) > 24.0\ \mathrm{mag\ arcsec^{-2}}$ (van Dokkum et al. 2015) for S\'{e}rsic model with $n=1$.} $\overline{\mu_e} > 24.4\ \mathrm{mag\ arcsec^{-2}}$ and effective radius $R_e > 1.5$ kpc, under the assumption that all matched LSBGs are physically associated with the hosts. Following these steps, we have identified $\sim$ 250 UDG candidates to be associated with Milky Way-like hosts in projection. They span a wide range of surface brightness, size, color, nucleation, and host properties, making this sample ideal for studying UDG spatial distributions, quenched fractions, mass-size relations, nucleation fractions, and their dependence on host properties.

@@

In practice, the measured central surface brightness might be biased by nuclear star clusters \citep{Neumayer2020,ELVES-II,Somalwar2020} or other contaminants. Therefore in this work, we use the average surface brightness within an effective radius $\sbeff$ to define UDGs. The difference between the average surface brightness and the central surface brightness can be analytically calculated for a \sersic{} profile: $\overline{\mu}_{\mathrm{eff}} - \mu_0 = 1.124$ for $n=1$ and 0.796 for $n=0.8$ \citep{Graham2005,Yagi2016}. Since dwarf galaxies and UDGs typically have \sersic{} index of $0.8 < n < 1.2$ \citep[e.g.,][]{vanDokkum2015,ELVES-I}, we take $\overline{\mu}_{\mathrm{eff}} - \mu_0 = 1$ as an average value to convert $\mu_0$ to $\mu_{\mathrm{eff}}$. In this work, we define UDGs to be galaxies with $r_e+\sigma(r_e) > 1.5$ kpc and $\sbeff + \sigma(\sbeff) > 25\ \sbunit$, where we also take the $1\sigma$ measurement errors into account. This definition maximizes the consistency with the definition in \citet{vanDokkum2015} while not losing UDGs harboring nuclear star clusters.

@@

We compare our UDG abundance with other surveys focused on MW analogs. In the SAGA survey \citep{SAGA-II}, 6 satellite galaxies (out of 127 spectroscopy-confirmed satellites around 36 MW analogs) satisfy the definition of UDG. Therefore the UDG abundance in \citet{SAGA-II} is about $N_{\rm UDG,\, SAGA} \approx 0.17\pm0.07$. In the ELVES survey \citep{CarlstenELVES2022}, there are 13 UDGs out of 358 satellites with secure distances ($P_{\rm sat} > 0.8$) in 30 MW analogs, leading to a UDG abundance of $N_{\rm UDG,\, ELVES} \approx 0.43\pm0.12$. \citet{Roman2017b} identified 11 UDGs around three galaxy groups in the IAC Stripe 82 Legacy Survey \citep{Fliri2016}. Because both SAGA and ELVES select satellites based on distance measurements and do not correct for completeness, we interpret their UDG abundances as lower limits. We note that \citet{Roman2017b} do not estimate the background contamination fraction and correct for completeness. We plot the UDG abundances of these surveys in the left panel of Figure \ref{fig:n_udg}, together with the results for larger galaxy groups and clusters \citep{Koda2015,Munoz2015,Roman2017a,Roman2017b,Janssens2017,vdBurg2017}. Among these studies, only \citet{vdBurg2017} applies a background subtraction and completeness correction to the raw counts.
% In the Auriga simulations, \citet{Liao2019} find $N_{\rm UDG,\,Auriga}=1.27\pm 1.06$ for UDGs in MW-mass hosts. 
The power-law regression result from \citet{vdBurg2017} $N_{\rm UDG} \propto M_h^{1.11\pm 0.07}$ is shown in gray. The UDG abundance of this work is highlighted as the red square. As shown in Figure \ref{fig:n_udg}, the scatter in the $N_{\rm UDG}-M_h$ relation gets larger at the lower halo mass end. Our UDG abundance is similar to the value in ELVES survey, but much higher than the value from SAGA. This might be explained by the different photometric depth of data used. As pointed out by \citet{CarlstenELVES2022,Font2022}, the SAGA survey reaches a surface brightness depth of $\sbeffr\approx 25\ \sbunit$, which is $\sim 2.5\ \sbunit$ shallower than the ELVES survey and this work. It is probable that SAGA missed a significant fraction of red and lower surface brightness galaxies. Our UDG abundance is marginally higher than the prediction from \citet{vdBurg2017} but is still consistent considering the large scatter of the power law.

@@

\subsection{A better definition of mass-size outliers}\label{sec:artifact}
% High and constant quenched fraction of UDGs is an artifact
In Section \ref{sec:sample_construction}, we have already argued that the hard size cut in UDG definition does not consider the mass-size relation, and we propose UPG which better represent the large-size tail of dwarf galaxy population. Here we discuss other defect in UDG definition which is the hard cut in surface brightness regardless of the galaxy color. We also discuss the advantages and disadvantages of different definitions of low surface brightness galaxies including UDG and UPG.

In Section \ref{sec:quench}, we find that the quenched fraction of UDGs is higher than the quenched fractions of UPGs and the satellite population of MW analogs. Furthermore, it almost remains a constant over $>1.5$ dex in stellar mass. We argue that UDG's high and constant quenched fraction is merely an artifact due to its contrived definition. We start from Figure \ref{fig:mass_size} where we show UDGs and UPGs on the mass-size plane color-coded by $\Delta(\mathrm{Quench}) = (g-i) - (g-i)_{Q}$ indicating how much a galaxy has been quenched (\S \ref{sec:quench}). The UDG sample includes a number of galaxies below the $1.5\sigma$ mass-size line, and these galaxies are mostly quenched (having zero or negative $\Delta(\mathrm{Quench})$). In order to help understand this phenomenon, we plot constant surface brightness lines of $\sbeff=25.0\ \sbunit$ for two different colors $g-i=0.4$ (blue solid line) and $g-i=0.8$ (red solid line) on the mass-size plane. For our redshift range, the cosmological dimming effect is negligible and the surface brightness is a constant with distance. For a given size and surface brightness (e.g., $\sbeff=25.0\ \sbunit$), redder galaxies have the same absolute magnitude as blue galaxies, but have higher stellar mass because they have higher mass-to-light ratio than blue galaxies. The color-$M_\star/L$ effect manifests itself on Figure \ref{fig:mass_size} by separating the two constant surface brightness lines. As a result, the UDG sample includes more red galaxies than blue ones due to the color-independent surface brightness cut at $\sbeff=25.0\ \sbunit$, and thus give rise to a high quenched fraction as we see in Figure \ref{fig:qfrac}. It is the definition of UDG that makes the quenched fraction high.

It worth emphasizing that the concept of UDG was originally proposed in a cluster context where the UDGs are mostly red and quenched \citep[e.g.,][]{vanDokkum2015} and the color-$M_\star/L$ effect does not matter that much. However, bluer UDGs are found in less dense environments. For a given stellar mass, the hard surface brightness cut in UDG definition removes blue galaxies because their surface brightnesses are higher but their $M_\star/L$ are lower than red ones. Therefore, the concept of UDG might not be optimal to study the mass-size outliers with different colors in different environments.
% We note that the color-$M_\star/L$ relation are only invoked when converting the absolute magnitude to stellar mass. Thus when studying the quenched fraction on size-luminosity plane (e.g., g-band absolute magnitude), the artifact would disappear \citep[e.g., see][]{Danieli2019}.

In this paper, we advocate an better definition for mass-size outliers: ultra-puffy galaxies (UPGs). They are defined to be $1.5\sigma$ above the average mass-size relation in the dwarf galaxy regime. The specific value above the average mass-size relation can vary to probe different parts of the size distribution. As we discussed in Section \ref{sec:sample}, a mass-dependent size cut for diffuse galaxies is more objective than a hard size cut because the average size gets larger with increasing stellar mass. However, the mass-size relation generally depends on the color, morphology, environment, and also redshift of the galaxies \citep[e.g.,][]{Graham2003,Trujillo2007,vanDokkum2013,Cappellari2013,Lange2015}. Fortunately, for dwarf galaxies ($10^{5.5}\ M_\odot < M_\star < 10^{8.5}\ M_\odot$) in the nearby Universe, \citet{ELVES-I} find that the mass-size relation is not sensitive to the color and morphology. This finding encourages definitions based on the mass-size relation for diffuse galaxies. From Figure \ref{fig:mass_size}, it is clear that the UPG definition naturally represents the large-size tail of the mass-size relation without introducing the artifacts due to the color-$M_\star/L$ relation. We further advocate studying the UPG population in both observation and simulation and explore different ``puffiness'' of UPGs (e.g., comparing UPG with $1.5\sigma$ with UPG with $2\sigma$).

>>>>

\author[0000-0002-8873-5065]{Peter Melchior}
\affiliation{Department of Astrophysical Sciences, 4 Ivy Lane, Princeton University, Princeton, NJ 08544, USA}

\author[0000-0002-1841-2252]{Rachael Beaton}
\affiliation{Department of Astrophysical Sciences, 4 Ivy Lane, Princeton University, Princeton, NJ 08544, USA}
\author[0000-0002-2991-9251]{Kirsten Casey}
\affiliation{Center for Cosmology and AstroParticle Physics (CCAPP), The Ohio State University, Columbus, OH 43210, USA}
\author[0000-0002-1841-2252]{Shany Danieli}
\affiliation{Department of Astrophysical Sciences, 4 Ivy Lane, Princeton University, Princeton, NJ 08544, USA}
\author[0000-0002-1841-2252]{Andy Goulding}
\affiliation{Department of Astrophysical Sciences, 4 Ivy Lane, Princeton University, Princeton, NJ 08544, USA}
\author[0000-0002-2704-5028]{Remy Joseph}
\affiliation{Department of Astrophysical Sciences, 4 Ivy Lane, Princeton University, Princeton, NJ 08544, USA}
\author[0000-0002-1841-2252]{Erin Kado-Fong}
\affiliation{Department of Astrophysical Sciences, 4 Ivy Lane, Princeton University, Princeton, NJ 08544, USA}

@@

%% Mark off the abstract in the ``abstract'' environment. 
\begin{abstract}
Large diffuse galaxies are hard to find, but understanding their environments and abundance, and ultimately their origins, is of intense interest. Using the Hyper Suprime-Cam Strategic Survey, we search for low surface-brightness galaxies and match them to Milky-Way like hosts in projection. We build a statistical sample of galaxies that are outliers in the mass-size relation and are satellites of Milky Way-mass galaxies in the nearby Universe ($0.01 < z < 0.04$). The number of these ``ultra-puffy'' galaxies per host is comparable to the observed abundance  at this halo mass in the Local Volume, and the galaxies follow an NFW profile around their halos. We argue that our definition, based on the mass-size relation, is more robust than the literature ultra-diffuse galaxy definition, which depends on surface brightness and thus yields different surface mass density cuts for quenched and star-forming galaxies. With our mass-size--based definition, we show that despite their extended size, the quenched fraction is comparable at a given satellite stellar mass between the ultra-puffy and normal dwarfs in MW-mass halo. These results hint that the size growth and quenching of mass-size outliers must happen separately. 
\end{abstract}

@@

\section{Introduction} \label{sec:intro}
In a variety of galaxy populations, an interesting subset of galaxies with low surface brightness and large physical size have been identified \citep[dubbed as low surface brightness galaxies or LSBGs, e.g.,][]{Sandage1984,Caldwell1987,Impey1988,McGaugh1995,Dalcanton1997a}. Recently, new excitement has been generated by the discovery of a large population of ultra-diffuse galaxies (UDGs) in the Coma cluster \citep{vanDokkum2015} , accompanied by many other works on detecting UDGs in clusters \citep[e.g.,][]{Koda2015,Mihos2015,Yagi2016,vdBurg2016,vdBurg2017,Lee2017,ManceraPina2018,Zaritsky2019}, groups \citep[e.g.,][]{Roman2017b,Greco2018,SAGA-II,CarlstenELVES2022}, and the field \citep[e.g.,][]{Leisman2017,Roman2019,Prole2019,Tanoglidis2021,Kadowaki2021}. There have been a variety of studies of these diffuse galaxy systems, ranging from their dark matter content \citep[e.g.,][]{Mowla2017,vanDokkum2018,vanDokkum2019,Wasserman2019,Keim2022}, to their globular cluster populations \citep[e.g.,][]{vanDokkum2017,Somalwar2020,Forbes2020,Danieli2022,Gannon2022}, and their stellar populations \citep[e.g.,][]{Gu2018,Ferre-Mateu2018,Villaume2022}. The UDGs are defined to be galaxies with large size ($r_e > 1.5$ kpc) and low surface brightness ($\sbcen > 24.0\ \sbunit$), and thus are very large in size for their stellar mass. Therefore, a central questions about LSBGs, including UDGs, is what causes them to be outliers from the mass-size relation.  
% Therefore, a central question about mass-size outliers including UDGs is what physical mechanisms are responsible for the formation and quenching of these large diffuse galaxies.

Different mechanisms are proposed to explain the presence of large diffuse galaxies in different environments. In the field, large size might be caused by star-formation feedback \citep{DiCintio2017,Chan2018}, early galaxy mergers \citep{Wright2021}, passing into and out of a more massive halo (so-called backsplash satellites, \citealt{Benavides2021}), or the galaxies inhabiting halos with higher spin \citep{Dalcanton1997,Amorisco2016,Liao2019}. In groups and clusters, the large size may ensue from tidal heating near pericenter \citep{Jiang2019} or via adiabatic expansion due to mass loss \citep{Tremmel2020}. Despite theoretical advances, it is still an open question how mass-size outliers like UDGs in certain environments are formed and quenched. It is also unclear whether mass-size outliers belong to a special class of galaxy or are just a natural extension of dwarf systems to lower surface brightness. To answer these questions, we can compare mass-size outliers with the ``normal'' satellites for which we have a better understanding.

Our best knowledge in the dwarf satellite regime is about the satellite system of our Milky Way (MW) and the Local Group (LG) \citep[e.g.,][]{McConnachie2012,Simon2019}. With the advent of deep sky surveys and dedicated spectroscopy programs, we are now mapping the satellite systems of MW analogs in the Local Volume and nearby Universe. The Satellites Around Galactic Analogs (SAGA) survey is designed to spectroscopically confirm the satellites of 100 MW analogs at $z\sim0.01$ \citep{SAGA-I,SAGA-II}. The Exploration of Local VolumE Satellites (ELVES) survey uses deep imaging data to identify satellite candidates of MW-like hosts in the Local Volume (LV, $D<12$ Mpc) and estimate surface brightness fluctuation distances from images to confirm satellites \citep{ELVES-I,ELVES-II,CarlstenELVES2022}. Having the MW satellites as a reference frame, we can study the formation and quenching of large diffuse galaxies by comparing them with the bulk of observed and simulated MW satellites. However, there are only $\sim 30$ confirmed UDGs associated with MW analogs in the literature \citep{Roman2017b,Cohen2018,SAGA-II,CarlstenELVES2022}. A larger sample of mass-size outliers (including UDGs) associated with MW hosts is needed to conduct such studies.

The goal of this paper is to study large diffuse galaxies in the reference frame of the satellite systems of MW analogs and try to elucidate their formation and quenching. For this purpose, we propose a new definition for mass-size outliers, dubbed as ``ultra-puffy galaxies'' (UPGs), based on the observed mass-size relation and its scatter. This definition is more robust to mass-to-light differences between red and blue galaxies compared with the UDG definition. Leveraging the depth and wide sky coverage of Hyper-Suprime Camera (HSC) imaging data, we are able to build a large statistical sample of mass-size outliers that are likely satellites of MW-mass hosts in the nearby Universe. The good statistics of our samples enable us to examine the abundance, size distribution, spatial distribution, and quenching of mass-size outliers. By comparing mass-size outliers with normal satellites and simulations, our results might shed light on the formation and evolution of the large diffuse galaxies.

@@

The layout of this paper is as follows. Section \ref{sec:data} describes the data used in this work. In Section \ref{sec:lsbg_search}, we describe the LSBG search in the HSC data, including source detection (\S \ref{sec:detection}), deblending (\S \ref{sec:deblending}), and modeling (\S \ref{sec:modeling}). In Section \ref{sec:sample_construction}, we propose a new definition for large and diffuse galaxies (the UPGs) and present our UDG and UPG samples. In Section \ref{sec:results}, we show statistical analysis for UDGs and UPGs including their abundances (\S \ref{sec:n_udg}), size distributions (\S \ref{sec:size_distr}), spatial distributions (\S \ref{sec:radial_distr}), and quenched fractions (\S \ref{sec:quench}). We argue why we need to move beyond UDG and discuss possible implications of our results on the formation and quenching of mass-size outliers in Section \ref{sec:discussion}. Section \ref{sec:summary} presents a summary of this work and prospects for the future. In this work, we refer to LSBGs as objects selected purely by surface brightness, whereas UDGs and UPGs are selected by both physical size and surface brightness. We adopt a flat $\Lambda$CDM cosmology from \citet{Planck15} with $\Omega_{\rm m}= 0.307$ and $H_0 = 67.7\ $km s$^{-1}$ Mpc$^{-1}$. We use the AB system \citep{Oke1983} for magnitudes. The stellar mass used in this work is based on a \citet{Chabrier2003} initial mass function.

@@

In practice, the measured central surface brightness might be biased by nuclear star clusters \citep{Neumayer2020,ELVES-II,Somalwar2020} or other contaminants. Therefore in this work, we use the average surface brightness within an effective radius $\sbeff$ to define UDGs. The difference between the average surface brightness and the central surface brightness can be analytically calculated for a \sersic{} profile: $\overline{\mu}_{\mathrm{eff}} - \mu_0 = 1.124$ for $n=1$ and 0.796 for $n=0.8$ \citep{Graham2005,Yagi2016}. Since dwarf galaxies and UDGs typically have \sersic{} index of $0.8 < n < 1.2$ \citep[e.g.,][]{vanDokkum2015,ELVES-I}, we take $\overline{\mu}_{\mathrm{eff}} - \mu_0 = 1$ as an average value to convert $\mu_0$ to $\mu_{\mathrm{eff}}$. In this work, we define UDGs to be galaxies with $r_e+\sigma(r_e) > 1.5$ kpc and $\sbeff + \sigma(\sbeff) > 25\ \sbunit$, where we also take the $1\sigma$ measurement errors into account. As a result, there are a few objects with smaller size but large uncertainty scattered in our UDG sample. This definition maximizes the consistency with the definition in \citet{vanDokkum2015} while not losing UDGs harboring nuclear star clusters.

@@

We compare our UDG abundance with other surveys focused on MW analogs. In the SAGA survey \citep{SAGA-II}, 6 satellite galaxies (out of 127 spectroscopy-confirmed satellites around 36 MW analogs) satisfy the definition of UDG. Therefore the UDG abundance in \citet{SAGA-II} is about $N_{\rm UDG,\, SAGA} \approx 0.17\pm0.07$. In the ELVES survey \citep{CarlstenELVES2022}, there are 13 UDGs out of 358 satellites with secure distances ($P_{\rm sat} > 0.8$) in 30 MW analogs, leading to a UDG abundance of $N_{\rm UDG,\, ELVES} \approx 0.43\pm0.12$. \citet{Roman2017b} identified 11 UDGs around three galaxy groups in the IAC Stripe 82 Legacy Survey \citep{Fliri2016}. Because both SAGA and ELVES select satellites based on distance measurements and do not correct for completeness, we interpret their UDG abundances as lower limits. We note that \citet{Roman2017b} do not estimate the background contamination fraction and correct for completeness. We plot the UDG abundances of these surveys in the left panel of Figure \ref{fig:n_udg}, together with the results for larger galaxy groups and clusters \citep{Koda2015,Munoz2015,Roman2017a,Roman2017b,Janssens2017,vdBurg2017}. Among these studies, only \citet{vdBurg2017} applies a background subtraction and completeness correction to the raw counts.
% In the Auriga simulations, \citet{Liao2019} find $N_{\rm UDG,\,Auriga}=1.27\pm 1.06$ for UDGs in MW-mass hosts. 
The power-law regression result from \citet{vdBurg2017} $N_{\rm UDG} \propto M_h^{1.11\pm 0.07}$ is shown in gray. The UDG abundance of this work is highlighted as the red square. As shown in Figure \ref{fig:n_udg}, the scatter in the $N_{\rm UDG}-M_h$ relation gets larger at the lower halo mass end, but this might be due to the small statistics in prior studies and the difference in completeness. Our UDG abundance is similar to the value in ELVES survey, but much higher than the value from SAGA. This might be explained by the different photometric depth of data used. As pointed out by \citet{CarlstenELVES2022,Font2022}, the SAGA survey reaches a surface brightness depth of $\sbeffr\approx 25\ \sbunit$, which is $\sim 2.5\ \sbunit$ shallower than the ELVES survey and this work. It is probable that SAGA missed a significant fraction of red and lower surface brightness galaxies. Our UDG abundance is marginally higher than the prediction from \citet{vdBurg2017} but is still consistent considering the large scatter of the power law.

@@

\subsection{Robust identification of mass-size outliers}\label{sec:artifact}
% In this paper, we propose a robust definition of mass-size outliers (the UPGs). Here we contrast UPGs with UDGs to elaborate the fact that UPGs better represent mass-size outliers.

% We also discuss the advantages and disadvantages of different definitions of low surface brightness galaxies including UDG and UPG.

In the left panel of Figure \ref{fig:mass_size}, we find that combining a hard physical size cut with a surface brightness cut carves out an interesting region of the mass-size plane that is complicated in two ways. On the one hand, there are many galaxies classified as UDGs that are not size outliers at $M_\star \sim 10^{7.5}\ M_\odot$. On the other hand, the hard surface brightness cut corresponds to very different surface mass density between red and blue galaxies, making it hard to directly compare red and blue diffuse populations. These issues manifest themselves in the high quenched fractions of UDGs. As seen in Figure \ref{fig:mass_size}, the UDG sample includes a number of galaxies below the $1.5\sigma$ mass-size line, and these galaxies are mostly quenched (having zero or negative $\Delta(\mathrm{Quench})$). In order to help understand this phenomenon, we plot constant surface brightness lines of $\sbeff=25.0\ \sbunit$ for two different colors $g-i=0.4$ (blue solid line) and $g-i=0.8$ (red solid line) on the mass-size plane. For our redshift range, the cosmological dimming effect is negligible and the surface brightness is a constant with distance. For a given size and surface brightness (e.g., $\sbeff=25.0\ \sbunit$), redder galaxies have the same absolute magnitude as blue galaxies, but have higher stellar mass because they have higher mass-to-light ratio than blue galaxies. As a result, the UDG sample includes more red galaxies than blue ones due to the color-independent surface brightness cut at $\sbeff=25.0\ \sbunit$, and thus give rise to a high quenched fraction as we see in Figure \ref{fig:qfrac}. It is the definition of UDG that makes the quenched fraction high.

% In Section \ref{sec:sample_construction}, we have already argued that the hard size cut in UDG definition does not map clearly onto the mass-size relation, and we propose identifying UPGs as outliers in the mass-size plane. Here we discuss another limitations in the UDG definition, namely the hard cut in surface brightness regardless of the galaxy color. We also discuss the advantages and disadvantages of different definitions of low surface brightness galaxies including UDG and UPG. 
% The above issues manifest themselves in the derived quenched fractions of UDGs. In Section \ref{sec:quench}, we find that the quenched fraction of UDGs is higher than the quenched fractions of UPGs and the satellite population of MW analogs. Furthermore, it almost remains a constant over $>1.5$ dex in stellar mass. 
% We argue that UDG's high and constant quenched fraction is merely an artifact of the definition. We start from Figure \ref{fig:mass_size}, where we show UDGs and UPGs on the mass-size plane color-coded by $\Delta(\mathrm{Quench}) = (g-i) - (g-i)_{Q}$ indicating how much a galaxy has been quenched (\S \ref{sec:quench}).

It is worth emphasizing that the concept of UDG was originally proposed in a cluster context where UDGs are mostly red and quenched \citep[e.g.,][]{vanDokkum2015} and the color-$M_\star/L$ effect does not matter that much. However, bluer UDGs are found in less dense environments. For a given stellar mass, the hard surface brightness cut in UDG definition removes blue galaxies because their surface brightnesses are higher but their $M_\star/L$ are lower than red ones. Therefore, the concept of UDG might not be optimal to study the mass-size outliers with different colors in different environments.
% We note that the color-$M_\star/L$ relation are only invoked when converting the absolute magnitude to stellar mass. Thus when studying the quenched fraction on size-luminosity plane (e.g., g-band absolute magnitude), the artifact would disappear \citep[e.g., see][]{Danieli2019}.

Our selection of UPGs is done directly based on the mass-size relation and thus alleviates the issues of UDG. They are defined to be $1.5\sigma$ above the average mass-size relation in the dwarf galaxy regime. The specific value above the average mass-size relation can vary to probe different parts of the size distribution. We further advocate studying the UPG population in both observations and simulations to explore different ``puffiness'' of UPGs (e.g., comparing UPG with $1.5\sigma$ with UPG with $2\sigma$).

However, in order to construct a UPG sample, one needs to know the distance, color, a color-$M_\star/L$ relation to convert observed magnitudes to stellar mass, and a mass-size relation and its scatter. None of these are easy to obtain and free from systematic errors. On the one hand, reliable mass-size relation and its scatter are not available for all mass ranges. \citet{ELVES-I} derive the mass-size relation below $M_\star \sim 10^{8.5}\ M_\odot$, whereas \citet{Lange2015} have relatively good constraints on the mass-size relation only above $M_\star \sim 10^{9}\ M_\odot$. The mass-size relation at $10^{8.5}\ M_\odot < M_\star < 10^{9}\ M_\odot$ is not well measured, nor is any dependence on color and morphology. In this work, we simply extrapolate the \citet{ELVES-I} mass-size relation to $10^9\ M_\odot$. At the same time, it is not clear whether the size distribution for a given mass can be well-described by a Gaussian distribution. Fig. 9 in \citet{ELVES-I} shows the residual in $\log(r_e)$ after fitting the mass-size relation, where the distribution of residuals is roughly Gaussian but skewed toward the large-size end. With enough statistics in the future, one might be able to characterize the shape of the large-size tail and define UPGs based on the percentiles of the size distribution. However, there are also other size definitions which help to reduce the scatter in the mass-size relation \citep[e.g.,][]{Miller2019,Trujillo2020} and can further benefit construct UPG samples.

In the right panel of Figure \ref{fig:mass_size}, we also plot constant average surface mass density lines, which are steeper than the mass-size relation but have the same slope as constant surface brightness lines. Hence another way to define diffuse galaxies is to select galaxies with low surface mass density. This method also alleviates the artifact introduced by color-$M_\star/L$ relation on the mass-size plane. We defer such studies to future works.
